# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

```{r}
unzip("activity.zip")
d<-read.csv("activity.csv",header=TRUE,sep=",")
```

## Total number of steps taken each day
```{r}
steps_per_day_sum <- aggregate(steps ~ date, d, sum)
hist(steps_per_day_sum$steps,col="red",main="Total Steps Taken Each Day",xlab="Steps")
```

### Mean and Median values of steps
```{r eval=FALSE}
mean(steps_per_day_sum$steps)
median(steps_per_day_sum$steps)
```

The mean value of the total number of steps taken each day is `r mean(steps_per_day_sum$steps)` and the median value is `r median(steps_per_day_sum$steps)`.


## Average daily activity pattern
```{r}
steps_per_interval_mean <- aggregate(steps ~ interval, d, mean)
plot(steps_per_interval_mean$interval,steps_per_interval_mean$steps,type="l",main="Steps, mean value",xlab="Interval",ylab="Steps, Mean")
```

### Interval with the maximum number of steps

```{r eval=FALSE}
max(steps_per_interval_mean$steps)                                  # maximum number of steps
steps_per_interval_mean[which.max(steps_per_interval_mean$steps),1] # corresponding interval
```

Maximum number of steps is `r round(max(steps_per_interval_mean$steps),2)`, from the interval `r steps_per_interval_mean[which.max(steps_per_interval_mean$steps),1]`.


## Imputing missing values
### Total number of rows with NA's

```{r}
sum(is.na(d))
```

There are `r sum(is.na(d))` in the given dataset. The chosen strategy to remove "NA" values from the dataset is to replace "NA" for steps with corresponding interval's mean value:

```{r}
library(plyr)
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm=TRUE))
d2 <-  ddply(d, ~ interval, transform, steps = impute.mean(steps))
```

Histogram for the new dataset:

```{r}
steps_per_day_clean_sum <- aggregate(steps ~ date, d2, sum)
hist(steps_per_day_clean_sum$steps,col="red",main="Total Steps Taken Each Day",xlab="Steps")
```

### Mean and Median values of steps
```{r eval=FALSE}
mean(steps_per_day_clean_sum$steps)
median(steps_per_day_clean_sum$steps)
```

The mean value of the total number of steps taken each day is `r mean(steps_per_day_clean_sum$steps)` and the median value is `r median(steps_per_day_clean_sum$steps)`.

There is no difference between the estimates from the first part of assignment. The chosen strategy has no impact on the computed mean and median values.

## Activity patterns at weekdays and weekends
```{r message=FALSE}
library(lubridate)
data_copy <- d2
data_copy[["weekdays"]]<-wday(data_copy$date)
workingdays<-subset(data_copy,weekdays!=6 & weekdays!=7)
weekend<-subset(data_copy,weekdays==6 | weekdays==7)
```

### Weekdays Activity
```{r}
workingdays_steps_per_interval_mean <- aggregate(steps ~ interval, workingdays, mean)
plot(workingdays_steps_per_interval_mean$interval,workingdays_steps_per_interval_mean$steps,type="l",main="Working Days Activity",xlab="Interval",ylab="Steps, Mean")
```

### Weekend Activity
```{r}
weekend_steps_per_interval_mean <- aggregate(steps ~ interval, weekend, mean)
plot(weekend_steps_per_interval_mean$interval,weekend_steps_per_interval_mean$steps,type="l",main="Weekend Activity",xlab="Interval",ylab="Steps, Mean")
```

On working days there is more activity in the earlier hours. The activity at the weekend days is more constant for all intervals.
